\LARGE \bf Detection of design pattern : A systematic mapping study

### 2016-01-01 article/main.tex

We decided to query only one database, Scopus \cite{c2}, that claims to be the
largest abstract and citation database of peer-reviewed literature.
It indexes the main Computer Science databases : Springer, IEEE, ACM and
others.
The main advantage of using Scopus is that it provides a functionnality to
export the result of a search query on its database.
Some of those exported results would not have been easily obtainable on some
publishers search engines that do not provide an export functionnality.
Consequently, Scopus was used mostly as a Search and Export portal over the
relevant publications, greatly reducing the effort needed to extract the
original articles set. 
% TODO: Original articles set <- better wording?!

From the two research questions we tried to maximise the number of publications
found on the subject.
% TODO: Thesaurus is a synonym dictionnary... therefore a "disjunction of
% thesaurus" makes no sense.
To achieve that purpose we used for the query a disjonction of thesaurus for
\textit{detection} on their respective contracted form : \textit{detect*},
\textit{recogni*} and \textit{ident*}.
We have then added to the query a disjonction of \textit{pattern} and \textit{motif}.
The query was limitied to the abstract, the title, and keywords of each article.

@@

The query resulted in a total of 403 articles.
Of those articles, we found two duplicates that were immediately removed from
the set of articles to analyse which brings down the number of articles for
screening to 401.

@@

The inter-rater agreement Cohen's kappa gave us the following result:
% TODO: citation needed

\begin{center}
  \begin{tabular}{ ccc }
    \bf Group & \bf \# of articles & \bf Cohen's kappa \\
    \hline
    1 & ??? & 0.88 \\
    2 & ??? & 0.66 \\
    3 & ??? & 0.91 \\
  \end{tabular}
\end{center}

@@

\begin{itemize}
  \item \textbf{Detection strategy}:
    This determine the means on which the detection is based.
    Possible values: Artificial intelligence, Logic/Formalisation, Graph,
    Metric, Automata, Instrumentation/profiling and Others.

  \item \textbf{Language dependance}:
    This category aims to classify detection methods on wether it is dependant
    on the language or not. 
    Possible values: Dependant or Independant
% TODO: This  specifies if the detection is langage dependent or independent,

@@

The yearly distribution gives us an interesting picture of the field.
Interest in design patterns detection technique started in the early
2000 and that interest is fading away since 2010.

@@

\begin{center}
\includegraphics[scale=0.7]{articles_by_years.png}
\end{center}

\subsection{Publication Venue Distribution}
% TODO: Need a graphic for article by publishers.

\begin{center}
  \small
  \begin{tabular}{ lll }
    \bf Category & \bf Class & \bf \# of articles \\
    \hline
    \multirow{7}{*}{Detection Srategy}
    & \multicolumn{1}{l}{AI}           & \multicolumn{1}{l}{11} \\
    & \multicolumn{1}{l}{Logic}        & \multicolumn{1}{l}{22} \\
    & \multicolumn{1}{l}{Graph}        & \multicolumn{1}{l}{15} \\
    & \multicolumn{1}{l}{Metric}       & \multicolumn{1}{l}{7} \\
    & \multicolumn{1}{l}{Automata}     & \multicolumn{1}{l}{1} \\
    & \multicolumn{1}{l}{Instr./Prof.} & \multicolumn{1}{l}{10} \\
    & \multicolumn{1}{l}{Pattern Mathching}
                                       & \multicolumn{1}{l}{8} \\
    & \multicolumn{1}{l}{Others}       & \multicolumn{1}{l}{23} \\
    \hline
    \multirow{2}{*}{\textbf{Language Dependance}}
    & \multicolumn{1}{l}{Dependant}    & \multicolumn{1}{l}{XXX} \\
    & \multicolumn{1}{l}{Independant}  & \multicolumn{1}{l}{XXX} \\
    \hline
    \multirow{3}{*}{\textbf{Analysis Type}}
    & \multicolumn{1}{l}{Static}       & \multicolumn{1}{l}{XXX} \\
    & \multicolumn{1}{l}{Dynamic}      & \multicolumn{1}{l}{XXX} \\
    & \multicolumn{1}{l}{Hybrid}       & \multicolumn{1}{l}{XXX} \\
    \hline
    \multirow{3}{*}{Detection Level}
    & \multicolumn{1}{l}{Execution}    & \multicolumn{1}{l}{8} \\
    & \multicolumn{1}{l}{Source}       & \multicolumn{1}{l}{44} \\
    & \multicolumn{1}{l}{Model}        & \multicolumn{1}{l}{7} \\
    \hline
    \multirow{2}{*}{Detection generality}
    & \multicolumn{1}{l}{Specific}     & \multicolumn{1}{l}{XXX} \\
    & \multicolumn{1}{l}{General}      & \multicolumn{1}{l}{XXX} \\
    \hline
    \multirow{4}{*}{Detected Pattern Types}
    & \multicolumn{1}{l}{Creationnal}  & \multicolumn{1}{l}{23} \\
    & \multicolumn{1}{l}{Structural}   & \multicolumn{1}{l}{34} \\
    & \multicolumn{1}{l}{Behaviour}    & \multicolumn{1}{l}{49} \\
    & \multicolumn{1}{l}{Other/Not Specified}
                                       & \multicolumn{1}{l}{17} \\
    \hline
    \multirow{6}{*}{Validation Method}
    & \multicolumn{1}{l}{Benchmark}    & \multicolumn{1}{l}{7} \\
    & \multicolumn{1}{l}{Case Study}   & \multicolumn{1}{l}{40} \\
    & \multicolumn{1}{l}{Experiment}   & \multicolumn{1}{l}{8} \\
    & \multicolumn{1}{l}{Survey}       & \multicolumn{1}{l}{0} \\
    & \multicolumn{1}{l}{Other}        & \multicolumn{1}{l}{1} \\
    & \multicolumn{1}{l}{None}         & \multicolumn{1}{l}{11}
  \end{tabular}
\end{center}

@@

Most of the detection strategies used fall either in the 
\textit{logic/formalisation} or \textit{other} class.
The first one was probably overly broad since it included any solution using
inference rules engines, formalisation and semantic web ontologies.
Surprisingly, the latter strategy was the focus of many articles.

@@

Design pattern detection can be done either statically or dynamically.

@@

The Gang of Four identified three design pattern types: creationnal,
behavioural and structural patterns.
Due to the nature of each pattern types, it is usally aknowledged that
structural patterns should be easier to detect than behavioural patterns and
that the later are easier than creationnal patterns (citation needed). 
It was expected that this ordering would appear in the number of articles
associated with each pattern types. 
However, we can see that the detection of behavioural patterns has been the
subject of more articles than the two others.

@@

This can be explained either by the fact that structural design patterns
detection is considered as an easy problem, thus less likely to be the main
focus of a research paper or by the possibility is that behavioural design
patterns are in fact easier to detect than structural ones.
However, we believe the first possibilty to be more likely.

@@

The majority of techniques are language dependant and the most used 
languages are Java and C++

But those detection methods can be easily converted to other languages
just by modifying the font-end.

@@

@@

Marjan Heričko and Simon Beloglavec[355] presents a way of identifying
composite design pattern, composite here corresponding to the composition of
atomic DP (MVC).

@@

@@

\subsection{Detected pattern types according to analysis type}
% Need graphics

@@

The distribution of articles between the reviewers may seem arbitrary in since it is not evenly distributed.
In fact one reviewer had two less articles to read.
This difference is due to an error when we compiled the articles that passed
the screening phase where one article was featured twice.
Since the remaining number of articles was not divisible by three, this same
reviewer had already one less article to read.
We believe this error has no effect on the overall result. \\

At first, we defined the detection level to be a disjoint classification category.
However during the classification phase of our study, we stumbled upon unfoseen cases
where the detection was done at two levels. We had to change the classification
scheme to allow those cases.

@@

So far, no other comparable studies has been done on the field of design pattern
recognition.

(article on method validation) have identified a problem in the evaluation of design
pattern detection methods. This problem has been observed in our evaluation of the
domain in that most articles evaluates their algorithms with case studies that are
hardly comparable.

>>>>

We decided to query only one database, Scopus \cite{c2}, that claims to be the
largest abstract and citation database of peer-reviewed literature.
It indexes the main Computer Science databases : Springer, IEEE, ACM and
others.
The main advantage of using Scopus is that it provides a functionality to
export the result of a search query on its database.
Some of those exported results would not have been easily obtainable on some
publishers search engines that do not provide an export functionality.
Consequently, Scopus was used mostly as a Search and Export portal over the
relevant publications, greatly reducing the effort needed to extract the
original articles set. 
% TODO: Original articles set <- better wording?!

From the two research questions we tried to maximize the number of publications
found on the subject.
% TODO: Thesaurus is a synonym dictionnary... therefore a "disjunction of
% thesaurus" makes no sense.
To achieve that purpose we used for the query a disjonction of thesaurus for
\textit{detection} on their respective contracted form : \textit{detect*},
\textit{recogni*} and \textit{ident*}.
We have then added to the query a disjonction of \textit{pattern} and \textit{motif}.
The query was limitied to the abstract, the title, and keywords of each article.

@@

The query resulted in a total of 403 articles.
Of those articles, we found two duplicates that were immediately removed from
the set of articles to analyze which brings down the number of articles for
screening to 401.

\begin{figure}
 \centering
  \begin{tabular}{ ccc }
   \bf Group & \bf \# of articles & \bf Cohen's kappa \\
   \hline {}
   1 & 133 & 0.88 \\
   2 & 133 & 0.66 \\
   3 & 134 & 0.91 \\
 \end{tabular}
 \caption{Cohen's Kappa coefficients}
\end{figure}

\begin{figure*}[tb]
 \centering
 \includegraphics[scale=0.2]{flow.jpg}
 \caption{Flow of information}
\end{figure*}

@@

The inter-rater agreement Cohen's kappa gave us the following result presented
in \textit{fig. 1}.

@@

\begin{itemize}
  \item \textbf{Detection strategy}:
    This determine the means on which the detection is based.
    Possible values: Artificial intelligence, Logic/Formalization, Graph,
    Metric, Automata, Instrumentation/profiling and Others.

  \item \textbf{Language dependence}:
    This category aims to classify detection methods on whether it is dependent
    on the language or not. 
    Possible values: Dependant or Independant
% TODO: This  specifies if the detection is langage dependent or independent,

@@

The yearly distribution gives us an interesting picture of the field.
Interest in design patterns detection technique started in the early
2000 and that interest plateaued between the years 2005 and 2013.
Since then the field lost steam with a number of publication per years
roughly equivalent to those at its inception.

@@

\begin{center}
 \begin{figure}
  \includegraphics[scale=0.450]{year_distribution.png}
  \caption{}
 \end{figure}
\end{center}

%\begin{center}
%\includegraphics[scale=0.7]{articles_by_years.png}
%\end{center}

\begin{center}
  \begin{figure}
   \small
   \begin{tabular}{ lll }
    \bf Category & \bf Class & \bf \# of articles \\
    \hline
    \multirow{7}{*}{Detection Strategy}
    & \multicolumn{1}{l}{AI}           & \multicolumn{1}{l}{11} \\
    & \multicolumn{1}{l}{Logic}        & \multicolumn{1}{l}{19} \\
    & \multicolumn{1}{l}{Graph}        & \multicolumn{1}{l}{13} \\
    & \multicolumn{1}{l}{Metric}       & \multicolumn{1}{l}{7} \\
    & \multicolumn{1}{l}{Automata}     & \multicolumn{1}{l}{1} \\
    & \multicolumn{1}{l}{Instr./Prof.} & \multicolumn{1}{l}{9} \\
    & \multicolumn{1}{l}{Pattern Matching}
                                       & \multicolumn{1}{l}{8} \\
    & \multicolumn{1}{l}{Others}       & \multicolumn{1}{l}{19} \\
    \hline
    \multirow{2}{*}{\textbf{Language Dependence}}
    & \multicolumn{1}{l}{Dependant}    & \multicolumn{1}{l}{37} \\
    & \multicolumn{1}{l}{Independent}  & \multicolumn{1}{l}{16} \\
    \hline
    \multirow{3}{*}{\textbf{Analysis Type}}
    & \multicolumn{1}{l}{Static}       & \multicolumn{1}{l}{25} \\
    & \multicolumn{1}{l}{Dynamic}      & \multicolumn{1}{l}{13} \\
    & \multicolumn{1}{l}{Hybrid}       & \multicolumn{1}{l}{17} \\
    \hline
    \multirow{3}{*}{Detection Level}
    & \multicolumn{1}{l}{Execution}    & \multicolumn{1}{l}{8} \\
    & \multicolumn{1}{l}{Source}       & \multicolumn{1}{l}{44} \\
    & \multicolumn{1}{l}{Model}        & \multicolumn{1}{l}{8} \\
    \hline
    \multirow{2}{*}{Detection generality}
    & \multicolumn{1}{l}{Specific}     & \multicolumn{1}{l}{40} \\
    & \multicolumn{1}{l}{General}      & \multicolumn{1}{l}{11} \\
    \hline
    \multirow{4}{*}{Detected Pattern Types}
    & \multicolumn{1}{l}{Creational}  & \multicolumn{1}{l}{22} \\
    & \multicolumn{1}{l}{Structural}   & \multicolumn{1}{l}{33} \\
    & \multicolumn{1}{l}{Behavior}    & \multicolumn{1}{l}{37} \\
    & \multicolumn{1}{l}{Other/Not Specified}
                                       & \multicolumn{1}{l}{12} \\
    \hline
    \multirow{6}{*}{Validation Method}
    & \multicolumn{1}{l}{Benchmark}    & \multicolumn{1}{l}{7} \\
    & \multicolumn{1}{l}{Case Study}   & \multicolumn{1}{l}{37} \\
    & \multicolumn{1}{l}{Experiment}   & \multicolumn{1}{l}{8} \\
    & \multicolumn{1}{l}{Survey}       & \multicolumn{1}{l}{0} \\
    & \multicolumn{1}{l}{Other}        & \multicolumn{1}{l}{1} \\
    & \multicolumn{1}{l}{None}         & \multicolumn{1}{l}{8}
   \end{tabular}
   \caption{Classification Results - Categories whose values represent disjoint sets are presented in bold}

  \end{figure}
\end{center}

\subsection{Publication Venue Distribution}
% Need a graphic for article by publishers.

@@

Most of the detection strategies used fall either in the 
\textit{logic/formalization} or \textit{other} class.
The first one was probably overly broad since it included any solution using
inference rules engines, formalization and semantic web ontologies.
Surprisingly, the latter strategy was the focus of many articles.

Another strong

@@

Design pattern detection can be done either statically or dynamically.
So

@@

The Gang of Four identified three design pattern types: creational,
behavioural and structural patterns.

Due to the nature of each pattern types, our intuition would dictate that
structural patterns should be easier to detect than behavioral patterns and
that the later are easier than creational patterns (citation needed). 
It was expected that this ordering would appear in the number of articles
associated with each pattern types. 
However, we can see that the detection of behavioral patterns has been the
subject of more articles than the two others.

@@

This can be explained either by the fact that structural design patterns
detection is considered as an easy problem, thus less likely to be the main
focus of a research paper or by the possibility that behavioral design
patterns are in fact easier to detect than structural ones.
We believe the first possibility to be more likely.

Of interest is the relation between the type of analysis and the type of
patterns detected. \textit{Fig. 5} and \textit{Fig. 6}
respectively presents the distribution of the number of articles and the
propotion.

\begin{figure}
 \centering
 \begin{tabular}{llll}
             & Static & Dynamic & Hybrid \\
 \hline
 Structural  & 18 & 7 & 8 \\
 Behavioural & 17 & 7 & 13 \\
 Creationnal & 14 & 1 & 7 \\
 Other       & 3 & 5 & 4 \\
 \end{tabular}
 \caption{Analysis type according to detected pattern type}
\end{figure}

\begin{figure*}[tb]
 \centering
 \includegraphics[scale=0.7]{analysis_v_pattern.png}
 \caption{Analysis type distribution per detected pattern type}
\end{figure*}

@@

The majority of techniques are language dependent and the most used 
languages are Java and C++, but those detection methods can be easily converted to other languages
just by modifying the font-end.

@@

Systems based on rules are examples of gendeals with an inference engine for which

@@

Heričko and Beloglavec[355] presents a way of identifying
composite design pattern, not to be confused with the definition of the 
composite pattern defined by the GoF, here composite pattern corresponds to the
composition of atomic design pattern (the Model View Controler being an
example of a composite pattern).

@@

Case studies are the most common methods of validation with a weak rate of
experiences and a total absence of true benchmarks.

Researchers in the Java world seem to have agreed on the use of JHotDraw,
JRefactory, JUnit, and AWT for evaluation.
However, while JHotDraw, a software known for its internal use of design
pattern, is the most frequently used, there is no clear benchmark.

Very few articles have taken the time to do a proper comparative evaluation.
De Lucia et al.[151] did an excellent work in that regard.
They compared their approach to existing tools and P-MARt's knowledge base
of pattern instances in JHotDraw and JRefactory, thus building a
strong case for their technique.
Unfortunately, most technique seem to be evaluated in a vacuum which makes it
difficult to compare their efficiency.

Also, because most technique are language dependant, comparative studies are
often times limited to case study in one language.
Again, few articles try to present results taken from software in written in
different languages.
This is problematic because no common ground for comparison exists between
technique whose evaluation is done on software written in different languages.
Gueheneux and Antoniol[255] presented an extensive evaluation of DeMIMA on
five open source Java projects and 33 industrial component written in C++.
While the use of closed source component is in itself problematic, the
evaluation of their system on both Java and C++ programs is noteworthy.

%TODO: citation needed
Other articles, due to the work involved in manually identifying instances of
design pattern for evaluation, have simply choosen to compare their result to
the output of existing tools.
Even if some use multiple tools to cross check the identified pattern instances,
we believe such evaluation methods are flawed since we don't know how well
those tools perform due to the absence of a standard benchmark.
The existence of pattern repository like P-MARt should supersede those
validation technique.

We have also noticed that there is no validation standards, there is a lot
of tools that can detect patterns, but since there are no standards 
none of those tools can be truly reliable.

All these issues seem to indicate a lack of maturity of the field of design
pattern detection.
At least, more work is needed compare the different techniques.

@@

@@

The distribution of articles between the reviewers may seem arbitrary in since it is not evenly distributed.
In fact one reviewer had two less articles to read.
This difference is due to an error when we compiled the articles that passed
the screening phase where one article was featured twice.
Since the remaining number of articles was not divisible by three, this same
reviewer had already one less article to read.
We believe this error has no effect on the overall result.

At first, we defined the detection level to be a disjoint classification
category.
However during the classification phase of our study, we stumbled upon
unforeseen cases where the detection was done at two levels.
We had to change the classification scheme to allow those cases.

@@

While no other systematic mapping study have been done on design pattern
detection technique, other reseachers attempted to address the problems
present in the field.

Pettersson, Lowe, and Nivre[190] identified a lack of proper evaluation of 
design pattern detection technique, in part due to the absence of an accepted
standard benchmark.
This problem has been observed in our evaluation of the domain in that most
articles evaluates their algorithms with case studies that are hardly
comparable.
They identified the existance of evaluation gold standard in JHotDraw, Swing
and AWT  and pushed for the definition of a standard benchmark made of varied
systems (language, size, etc.)
